Community-Conditioned Language Models

### 2020-11-20 paper/paper.tex
todos discussion

We experiment with two kinds of model architecture: LSTM
\citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm? Even If
  it's simple it may be worth specifying due to the prevalence of
  bilstm everywhere.}  and Transformer \citep{Vaswani2017}.  In either
case, the model is organised as a standard $n$-layer neural sequence
encoder (we set $n=3$ for all models). As is standard, word tokens are
first converted to vectors using a trainable embedding layer. At the
other end, word tokens are predicted using a softmax projection layer. The set of word tokens is a
pre-determined vocabulary.  These encoders form the core of the
community-conditioned language models (CCLM), and we call them
unconditioned models.

@@

For the LSTM, we trained the models on a simple left-to-right language modeling task
with cross-entropy loss over the vocabulary. 
Because the transformer operates on all tokens in the sequence at once,
the language modeling was achieved by masking and incrementally un-masking input tokens
in a left-to-right fashion, following \citet{BERT}\jp{Or some other standard paper}.  
We used the AdamW \citep{Loshchilov2019} optimization algorithm, 
with an initial learning rate of \num{0.001}.\jp{Also for the transformer? No simulated annealing?}
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{This should make no difference (but I'm not sure on how you are taking averages; see other comment.)}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more general interest topics,
such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
range of people.
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and other linguistic patterns.

@@

We consider the average information gain of each CCLM over its baseline
un-conditioned counterpart with the same sequence encoder architecture.
For a given message, the information gain is defined as the ratio of the
perplexity of the baseline model to that of the CCLM:
\[
  IG_{CCLM}(m) = \frac{Ppl_{LM}(m)}{Ppl_{CCLM}(m)}
\]
\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
  then taking the average over all messages:
  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m))$. Taking the exponential
  gives a ratio of perplexity.  Here it seems that you have commuted
  exponential and average. It is rather weird to do average in the
  exponential space without normalising to a probability.}
\begin{table*}
  \centering
  % \input{floats/info_gain.tex}
  \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
  \label{tab:model-info-gain}
\end{table*}

@@

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them from the message-community co-occurence matrix,
using data which spans January 2014 to April 2017. We refer the reader
to \citet{Kumar2018} % JP: I actually cannot find in their paper what they do.
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether and how much a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data. 
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

@@

We find a weak correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations. However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
CCLM and the indiscernibility of the corresponding community
\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
and only large communities in our test set, it is difficult to
provide a conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. However, even the most complex communities
(perplexity per word above 100) have an indiscernibility of about 3
(out of a maximum of 46), so they retain fairly specific patterns:
they are not simply an even mixture of the kind of comments found in
other communities.

>>>>

We experiment with two kinds of model architecture: a simple unidirectional LSTM
\citep{Hochreiter1997} and a Transformer \citep{Vaswani2017}.  In either
case, the model is organised as a standard $n$-layer neural sequence
encoder (we set $n=3$ for all models). As is standard, word tokens are
first converted to vectors using a trainable embedding layer. At the
other end, word tokens are predicted using a softmax projection layer. The set of word tokens is a
pre-determined vocabulary.  These encoders form the core of the
community-conditioned language models (CCLM), and we call them
unconditioned models.

@@

For the LSTM, we trained the models on a simple left-to-right language modeling task
with cross-entropy loss over the vocabulary. 
Because the transformer operates on all tokens in the sequence at once,
the language modeling was achieved by masking and incrementally un-masking input tokens
in a left-to-right fashion (so we have an auto-regressive\jp{check} model).
We used the AdamW \citep{Loshchilov2019} optimization algorithm,
with an initial learning rate of \num{0.001}, without any extra control on the decay of learning rate.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).
\bn{talk about this in terms of mutual information instead!}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more general interest topics,
such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
range of people.
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and other linguistic patterns.

@@

We consider the average information gain of each CCLM over its baseline
un-conditioned counterpart with the same sequence encoder architecture.
For a given message, the information gain is defined as the ratio of the
perplexity of the baseline model to that of the CCLM:
\[
  IG_{CCLM}(m) = \frac{Ppl_{LM}(m)}{Ppl_{CCLM}(m)}
\]
\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
  then taking the average over all messages:
  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
  gives a ratio of perplexity.  Here it seems that you have commuted
  exponential and average. It is rather weird to do average in the
  exponential space without normalising to a probability.

  Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.

}
\begin{table*}
  \centering
  % \input{floats/info_gain.tex}
  \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
  \label{tab:model-info-gain}
\end{table*}

@@

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them from the message-community co-occurence matrix,
using data which spans January 2014 to April 2017. We refer the reader
to \citet{Kumar2018}
\jp{What they say they do (sec 6.2) More formally, we train user and community embeddings basedon the bipartite multigraph between users and communities, whereone user-community edge,(ui,cj), exists in the graph for each timea user,ui, posts on community,cj. Adapting the approach used ina number of recent successful works [27,53,62], we learn user andcommunity embeddings,ui∈Rd,cj∈Rd, from this graph using anegative-sampling optimization algorithm
}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether and how much a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data. 
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

@@

We cannot exclude a correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the
% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations. However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
CCLM and the indiscernibility of the corresponding community
\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
and only large communities in our test set, it is difficult to
provide a conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. However, even the most complex communities
(perplexity per word above 100) have an indiscernibility of about 3
(out of a maximum of 46), so they retain fairly specific patterns:
they are not simply an even mixture of the kind of comments found in
other communities.
